Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-07-04 WRITEUP/ms.tex
added quick summary of theoretical work by Behrendt et al on clump structure.
added some discussion of virial relation and Martin & Kennicutt 2001 cite to Toomre

Earlier epochs still represent an essentially uncharted territory for ISM investigations. At present, it remains unclear how \SF proceeds in the (sub-)$L^*$ galaxy population at \z$\gtrsim$\,6  which is responsible for producing the bulk of the UV photons that reionized the Universe.
%
While ALMA has enabled the detection of \cii158\,$\micron$ and CO line emission in normal (SFR$<$\,100\,\Msun\,yr\pmOne) galaxies at \z$>$\,6 over the past few years \citep[e.g.,][]{Carniani18b, Odorico18a}, the first spatially resolved observations are just starting to become available \citep[e.g., ][]{Jones17a}.

@@

\section{Physical Background}\label{sec:Back}
We begin by introducing some basic equations concerning the empirical relations commonly mentioned in observational studies of molecular clouds in the literature and the gravitational instability of galactic disks, which might be driving MCC formation. Such notions will be used in the subsequent analysis of our simulations.

\subsection{Larson's Relations and Virial Parameter}\label{sec:PVE}
\citet{Larson81a} discussed a number of relations among Galactic
molecular cloud
properties, namely the linewidth-size, density-size, and mass-size relations. Larson relations are routinely used for comparing properties of molecular structures in different galactic environments. They also represent a useful framework to analyze our results as they
have been argued to arise from the interplay between gravity and turbulence encoded in the virial theorem
(note, however, the alternative interpretation involving gravitational collapse by
e.g., \citealt{Ballesteros-Paredes11a}).

@@

Based on Equation~\ref{eqn:SVE}, a one-to-one mapping between $\sigma_{\rm gas}^2/R$ and $\Sigma$ is therefore expected for a virialized cloud, since $\sigma_{\rm gas}^2/R\propto\Sigma$. Deviations from this relation are usually attributed to a significant contribution from external pressure as per Equation~\ref{eqn:v0} (see e.g., \citealt{Heyer09a, Hughes10a, Hughes13b, Meidt13a}). Similar conclusions are also reached from the analysis of clouds forming in Milky-Way like galaxy simulated with a $\simeq 4\,{\rm pc}$ resolution \citep[i.e.][]{grisdale:2018}.

@@

In cases where stellar component plays an important role in the dynamics, the virial parameter becomes
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G (M_{\rm gas} + M_{\star})} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}
where $M_\star$ is the stellar mass within the MCC volume.

@@

The onset of gravitational instability is tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. For axisymmetric modes, the dispersion relation for the growth of density perturbations in a rotating, turbulent disk of finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma_{\rm disk}^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{\altomega}\frac{d}{d\altomega}\left(\altomega^2\Omega\right)\,.
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
In Equation~\ref{eqn:3Ddisp} the terms on the right hand side are related to rotation, self-gravity and internal pressure, respectively. Heuristically, the instability can be understood by considering the scale at which gravitational potential overcomes the internal energy. Gravity dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, differential rotation in disk galaxies can stabilize perturbations that might otherwise collapse for $L > L_{\rm rot}$, where $L_{\rm rot}$ is set by $\kappa$. As a result, disks are unstable to gravitational collapse on scales between $L_J < L < L_{\rm rot}$.

>>>>

Earlier epochs still represent an essentially uncharted territory for ISM investigations. At present, it remains unclear how \SF proceeds in the (sub-)$L^*$ galaxy population at \z$\gtrsim$\,6  which is responsible for producing the bulk of the UV photons that reionized the Universe.
%
While ALMA has enabled the detection of \cii158\,$\micron$ and CO line
emission in normal (SFR$<$\,100\,\Msun\,yr\pmOne) galaxies at \z$>$\,6
over the past few years \citep[e.g.,][]{Carniani18b, Odorico18a}, the
first spatially resolved observations are just starting to become
available \citep[e.g., ][]{Jones17a}.
% mm2
    On the theoretical side, a schematic investigation by
    \citet{Behrendt16} revealed that kiloparsec-scale clumps likely consist
    of many smaller clumps formed by gravitational instability, a hierarchical
    structure that persists in an idealized disk model by \citet{Behrendt19}.

@@

\section{Physical Background}\label{sec:Back}
%mm2 [revised paragraph]
We begin by introducing some empirical relations commonly mentioned in
observational studies of molecular clouds in the literature, as well
as describing the gravitational instability of galactic disks, which
might be driving MCC formation. These concepts will be used in the subsequent analysis of our simulations.

\subsection{Larson's Relations and Virial Parameter}\label{sec:PVE}
\citet{Larson81a} discussed a number of relations among Galactic
molecular cloud
properties, namely the linewidth-size, density-size, and mass-size relations. Larson relations are routinely used for comparing properties of molecular structures in different galactic environments. They also represent a useful framework to analyze our results as they
have been argued to arise from the interplay between gravity and
% mm2 turbulence encoded in the
    velocity dispersion given by
virial theorem
(note, however, the alternative interpretation involving gravitational collapse by
e.g., \citealt{Ballesteros-Paredes11a}).

@@

Based on Equation~\ref{eqn:SVE}, a one-to-one mapping between
$\sigma_{\rm gas}^2/R$ and $\Sigma$ is therefore expected for a
virialized cloud, since $\sigma_{\rm
  gas}^2/R\propto\Sigma$.
% mm2
    \citep{Heyer09a} pointed out that the original size-linewidth
    relation implies constant surface density in this interpretation.
    The limited dynamic range in column density of CO observations
    appears to account for this \citep[see also][]{Ballesteros-Paredes11a}.
Deviations from this relation are
% mm2 usually
        often
attributed to a significant contribution from external pressure as per
Equation~\ref{eqn:v0} (see e.g., \citealt{Heyer09a, Hughes10a,
  Hughes13b, Meidt13a}). Similar conclusions are also reached from the
analysis of clouds forming in a Milky-Way like galaxy simulated with a $\simeq 4\,{\rm pc}$ resolution \citep[i.e.][]{grisdale:2018}.

@@

In cases where a stellar component plays an important role in the
dynamics, the virial parameter becomes
%mm2 [citation or derivation needed?]   ***********
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G (M_{\rm gas} + M_{\star})} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}
where $M_\star$ is the stellar mass within the MCC volume.

@@

The onset of gravitational instability is tightly connected to \SF
%mm2 [added Martin & Kennicutt 2001]
\citep[e.g.,][]{Kennicutt89a, Martin01, Wang94a, Li05b, Li06a}. For axisymmetric modes, the dispersion relation for the growth of density perturbations in a rotating, turbulent disk of finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma_{\rm disk}^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{\altomega}\frac{d}{d\altomega}\left(\altomega^2\Omega\right)\,.
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
In Equation~\ref{eqn:3Ddisp} the terms on the right hand side are related to rotation, self-gravity and internal pressure, respectively. Heuristically, the instability can be understood by considering the scale at which gravitational potential overcomes the internal energy. Gravity dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, differential rotation in disk galaxies can stabilize perturbations that might otherwise collapse for $L > L_{\rm rot}$, where $L_{\rm rot}$ is set by $\kappa$. As a result, disks are unstable to gravitational collapse on scales between $L_J < L < L_{\rm rot}$.
